---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF - SS 2022/23"
author: "Tom Jodl 12598878"
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r include=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(readr)
library(dplyr)
library(knitr)
library(ggplot2)
library(remotes)
library(margins)
```



## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers
    (Martrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your
    results.
-   Use a level 2 heading for each new exercise and answer each subtask
    next to it's bullet point or use a new level 3 heading if you want.
-   Always render the R code for your solutions and make sure to include
    the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless
        specifically instructed to).
-   Always submit both the rendered document(s) as well as your source
    Rmarkdown document. Submit the files separately on moodle, **not**
    as a zip archive.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

### a)

Start by initializing a new R project with git support, called
    `exeRcise-sheet-3`. If you forgot how to do this, you can follow
    this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
    
### b)

Commit the files generated by Rstudio.

### c)

For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

### d)

Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
commit after this answer, see *1c*)

#### Positiv:
Distributed Version Control

Branching and Merging

#### Negativ:
Learning Curve

Larger Datasets
    
### e)

Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves.
    
## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

### a)

Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:[github](https://github.com/TomJodl/exeRcise-sheet-3-neu)
    
### b)

Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
    
### c)

Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

## Exercise 3: Baby-Names in Munich (4.5 Points)

Download the latest open datasets on given names ("Vornamen") from the
open data repository of the city of Munich for the years `2022` and
`2021`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

### a)

Download the data for both years and track it in git. For small
    datasets like these adding them to git is not a problem.

### b)

Load the data for both years into R. Check the type of the count
    variable ("Anzahl") and look into the data to determine why it is
    not numeric? Fix the problem in an appropriate manner, it is OK if
    some of the counts are inaccurate because of this. Explain your
    solution and the repercussions.

#### Loading the data   

```{r}
# Load data for 2021
data_2021 <- read.csv("vornamen_2021.csv")

# Load data for 2022
data_2022 <- read.csv("open_data_portal_2022.csv")
```

#### Check the Variable

```{r echo=FALSE}
str(data_2021$Anzahl)

str(data_2022$Anzahl)
```

#### Making "Anzahl" Numeric

```{r}
# Fix the "Anzahl" variable in 2021 data
data_2021$Anzahl <- as.numeric(gsub("[^0-9]", "", data_2021$Anzahl))

# Fix the "Anzahl" variable in 2022 data
data_2022$Anzahl <- as.numeric(gsub("[^0-9]", "", data_2022$Anzahl))

```

After fixing the "Anzahl" variable, be aware that some counts may become inaccurate due to the removal of non-numeric characters. Missing values or non-numeric entries will be converted to NA.

### c)

Calculate the total number of babies born in Munich in 2022
    and 2021. Which year had the bigger baby-boom?
    
#### Calculate total number of babies
    
```{r}
# Calculate the total number of babies born in 2022
total_babies_2022 <- sum(data_2022$Anzahl)

# Calculate the total number of babies born in 2021
total_babies_2021 <- sum(data_2021$Anzahl)
```

#### Compare the numbers

```{r}
if (total_babies_2022 > total_babies_2021) {
  bigger_baby_boom <- 2022
} else if (total_babies_2022 < total_babies_2021) {
  bigger_baby_boom <- 2021
} else {
  bigger_baby_boom <- "Both years had an equal number of babies born"
}
bigger_baby_boom
```

### d)

Add a new column `year` to both datasets which holds the correct
    year for each.
    
```{r}
# Add the 'year' column to the 2021 dataset
data_2021$year <- 2021

# Add the 'year' column to the 2022 dataset
data_2022$year <- 2022
```

### e)

Combine both datasets into one using `bind_rows()`.

```{r}
combined_data <- bind_rows(data_2021, data_2022)
```

### f)

Combine the counts for same names to determine the most popular
    names across both years. Print out the top 10 names in a nicely
    formatted table for both years. Include a table caption.
    
```{r}
# Group the data by name and calculate the total count
name_counts <- combined_data %>%
  group_by(Vorname) %>%
  summarize(Total_Count = sum(Anzahl)) %>%
  arrange(desc(Total_Count))

# Print the top 10 names in a formatted table
top_10_names <- head(name_counts, 10)
kable(top_10_names, caption = "Top 10 Most Popular Names")
```

## Exercise 4: Chat GPT + apply (3 points)

For this task: Specifically use ChatGPT to solve the task and submit
your prompts in addition to the solution

### a)  

The code below does not work because the wrong apply function has
    been used. Find out which apply function would be correct and why it
    did not work. Correct the code. Also calculate the rowwise means.

#### Create a sample data frame

tax_data \<- data.frame( Name = c("Munich GmbH", "ABC Inc.", "Backpacks
1980", "Bavarian Circus"), Tax_2019 = c(5000, 4000, 6000, 3500),
Tax_2020 = c(4800, 4200, 5800, 3700), Tax_2021 = c(5200, 3800, 5900,
3400) )

#### Calculate column-wise means

column_means \<- lapply(tax_data\ [,-1\], 2, mean)

column_means

#### Prompt:

Task

#### Solution:

The code provided is attempting to calculate the column-wise means of the numeric columns in the tax_data data frame using the lapply() function. However, the second argument of lapply() is incorrect and should be replaced with the correct apply function.

The correct apply function to use in this case is sapply() because it simplifies the output into a vector or matrix, which is suitable for calculating column-wise means.

Here's the corrected code to calculate the column-wise means and row-wise means:

```{r}
# Create a sample data frame
tax_data <- data.frame(
  Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
  Tax_2019 = c(5000, 4000, 6000, 3500),
  Tax_2020 = c(4800, 4200, 5800, 3700),
  Tax_2021 = c(5200, 3800, 5900, 3400)
)

# Calculate column-wise means
column_means <- sapply(tax_data[, -1], mean)

# Calculate row-wise means
row_means <- rowMeans(tax_data[, -1])

# Print the column-wise means
column_means
# Print the row-wise means
row_means
```

After executing the corrected code, column_means will contain the column-wise means for the numeric columns (Tax_2019, Tax_2020, Tax_2021), and row_means will contain the row-wise means for each row of the numeric columns.

Please note that in the corrected code, sapply() is used instead of lapply() to calculate the column-wise means. Additionally, rowMeans() is used to calculate the row-wise means.

### b)

Using ChatGPT try to understand what the rapply() function does.
    Create an easy example with mock data where the function is used and
    explain it in your words.
    
#### Data:

```{r}
directory <- list(
  file1 = "file1.txt",
  subdirectory1 = list(
    file2 = "file2neu.txt",
    file3 = "file3.txt"
  ),
  subdirectory2 = list(
    file4 = "file4.txt"
  )
)
```

```{r}
char_count <- rapply(directory, function(x) nchar(x), how = "list")
```

```{r}
char_count$file1

char_count$subdirectory1$file2
```

rapply() is similar to the lapply() function. It applies a specified function to each element of a list, and if any of the elements are themselves lists, it recursively applies the function to those nested lists as well. rapply() looks through all the nested structures before it gives out a result. It can also simplefy resulst by making them a vector, matrix or array. Or like in the example into a list. 
In summary the lapply function can be used when the desired result is a list and the rapply fuction can be used when the desired result is a result.